\documentclass[master.tex]{subfiles}

\chapter{Evaluation Plan}

Evaluation will split into 2 parts as the following

\section{Users Experience}

As phometa is implementation based project. One way to get a truthful result is to do a survey. I will recruit several volunteers (mainly as other students) who interest logic to do the following

\begin{enumerate}
    \item do the tutorial that I have prepared in advance. In the tutorial, it will teach volunteer how to prove a theorem using traditional derivation tree on pen-and-paper then teach how to do the same thing with phometa. So they will know all of necessary background.
    \item prove 6 easy-to-prove propositions of first-order logic, measure the time taken for each question. Half of questions will be proved via pen-and-paper method then proved via phometa. The other half are vice versa. So there is no bias between pen-and-paper and phometa.
    \item create one formal system of their choice, and prove something with it.
    \item do a questionnaire by rating 1 (Strongly disagree) to 5 (Strongly agree) on the following statement
    \begin{itemize}
         \item Phometa is easier use than pen-and-paper method.
         \item Phometa take less time to prove theorem than pen-and-paper method.
         \item Phometa is responsive-in-time enough to be interactive proof assistant.
         \item I enjoy proving theorem on phometa.
         \item I enjoy create a new formal system on phometa.
         \item If I have a new idea related to formal system. I will use phometa to clarify it.
         \item If I have know someone who interest on this area, I will recommend them to try phometa.
         \item If I have time, I wish to contribute on phometa project.
    \end{itemize}
\end{enumerate}

The entire survey for each volunteer should take 1 or 2 hours depends on their prior knowledge on this topic. I believe that it is not that hard to find a people to do this survey since this would be fun enough to keep concentration for 1 or 2 hours. Moveover, I will also mansion all of volunteers names (if they not say otherwise) as an honourable contribution on evaluation chapter.

\section{Experts Feedback}

I will sent phometa to a few people who expert on proof assistant and formal system to comment. This could give another aspect of evaluation from users experience. If both of parts has positive outcome, this means phometa succeed as a proof assistant.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "master"
%%% End:
